Python as a machine learning language: experiences from the scikit-learn project

### 2013-06-17 paper.tex
wording

% \subsection{Cython for performance}

Our implementation guidelines emphasize on writing short and readable code,
easily maintainable by external contributors, while being as efficient as
possible.

Whenever practicable, algorithms implemented in \sklearn are thus
directly written in pure Python. For intensive numerical computations though,
the library mostly relies on NumPy vector operations, which allows for the code
to remain concise, readable and efficient. For all algorithms that cannot be
easily and efficiently expressed as NumPy operations, we rely on Cython
\citep{behnel2011cython}  for writing the critical parts and achieving decent
performance and scalability. Cython is a compiled programming language that
extends Python with static typing. It produces efficient C extension modules
that are directly importable from the Python run-time system. Examples of
algorithms written in Cython include stochastic gradient descent for linear
models, some graph-based clustering algorithms or decision trees.  A problem
arising from the use of Cython, however, is that passing pure Python functions
as argument, while possible, degrades performance significantly due to the
run-time checks needed for each function invocation. This is especially
problematic for example for kernel methods, which would benefit from accepting a
kernel function as argument. A potential solution to this problem would be for
\sklearn to provide a Cython-level API, thereby allowing the user to pass
Cython-implemented kernel functions.

To facilitate the adoption of the library, we also try to limit to a bare
minimum the set of external dependencies. As such, NumPy and SciPy -- the two
most common dependencies in the Python scientific ecosystem -- are actually the
only two required dependencies for the library to fully work. Some modules may
rely on other external libraries, but in such cases we tend to prefer to
directly ship them within the codebase rather than adding new dependencies for
the user to install. In particular, the codebase includes modified versions of
the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using
Cython modules.

>>>>

Our implementation guidelines emphasize on writing efficient but readable code.
In particular, we focus on making the codebase easily maintainable and
understandable in order to favour external contributions. Whenever practicable,
algorithms implemented in \sklearn are thus directly written in pure Python. For
intensive numerical computations though, the library mostly relies on NumPy
vector operations, which allows for the code to remain concise, readable and
efficient. For all algorithms that cannot be easily and efficiently expressed as
NumPy operations, we rely on Cython \citep{behnel2011cython}  for writing the
critical parts and achieving decent performance and scalability. Cython is a
compiled programming language that extends Python with static typing. It
produces efficient C extension modules that are directly importable from the
Python run-time system. Examples of algorithms written in Cython include
stochastic gradient descent for linear models, some graph-based clustering
algorithms or decision trees.  A problem arising from the use of Cython,
however, is that passing pure Python functions as argument, while possible,
degrades performance significantly due to the run-time checks needed for each
function invocation. This is especially problematic for example for kernel
methods, which would benefit from accepting a kernel function as argument. A
potential solution to this problem would be for \sklearn to provide a Cython-
level API, thereby allowing the user to pass Cython-implemented kernel
functions.

To facilitate the adoption and the distribution of the library, we also try to
limit to a bare minimum the set of external dependencies. As such, NumPy and
SciPy -- the two most common dependencies in the Python scientific ecosystem --
are actually the only two required dependencies for the library to fully work.
Some modules may rely on other external libraries, but in such cases we tend to
prefer to directly ship them within the codebase rather than adding new
dependencies for the user to install. In particular, the codebase includes
modified versions of the \textsf{LIBSVM} and \textsf{LIBLINEAR} libraries
\citep{chang2011libsvm,fan2008}, both written in C{}\verb!++! and wrapped using
Cython modules.
